---
title: "class13"
author: "Jiaye Chen"
date: "February 19, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In silico docking of drugs to HIV-1 protease

### Prepare HIV-protease molecule for docking

First get a PDB structure from the online database then clean it up for docking
```{r}
library(bio3d)
hsg.file <- get.pdb("1hsg")
hiv <- read.pdb(hsg.file)
hiv
```

Read file into R ad clean it up (remove the non-protein bits) and also we will save the ligand in to a separate file
```{r}
hiv <- read.pdb(hsg.file)
hiv
```

```{r}
prot <- trim.pdb(hiv, "protein")
lig <- trim.pdb(hiv, "ligand")
```

```{r}
write.pdb(prot, file="1hsg_protein.pdb")
write.pdb(lig, file="1hsg_ligand.pdb")
hsg_protein <- read.pdb("1hsg_protein.pdb")
hsg_ligand <- read.pdb("1hsg_ligand.pdb")
show(hsg_protein)
show(hsg_ligand)
```

```{r}

```

Q1: What is the name of the two non protein resid values in this structure? What does resid
correspond to and how would you get a listing of all reside values in this structure? 
```{r}
# Ans: HOH (127), MK1 (1) 
```

### 2.3 Inspecting your docking results
```{r}
library(bio3d)
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```
```{r}
res <- read.pdb("all.pdbqt", multi=TRUE)
ori <- read.pdb("ligand.pdbqt")
rmsd(ori, res)
```


## Normal Mode Analysis
```{r}
pdb <- read.pdb("1HEL")

m <- nma(pdb)
```
```{r}
plot(m)
```
```{r}
mktrj(m,mode = 7, file = "nma_7.pdb")
```


