---
title: "Class_06"
author: "Jiaye Chen"
date: "January 24, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section 1: Reading files again :-)

Here we are going to try using the **read.table()** function and friends to read some example flat files.

```{r}
plot(1:10)
```

Back to file reading...
```{r}
File1 <- ("https://bioboot.github.io/bimm143_W19/class-material/test1.txt")
data1 <- read.csv(File1)
data1
```

Reading text2
```{r}
File2 <- ("https://bioboot.github.io/bimm143_W19/class-material/test2.txt")
data2 <- read.csv(File2, sep = "$")
data2
```

Reading text3
```{r}
File3 <- ("https://bioboot.github.io/bimm143_W19/class-material/test3.txt")
data3 <- read.table(File3, header = FALSE)
data3
```

## Section 2. R Functions

My first silly function

```{r}
add <- function(x, y = 1) {
  # Sum the input x and y
  x+y
}
```

Lets try using this function
```{r}
add(1)
add(3,5)
add(c(1,2,3),4)
```

Function "rescale"
```{r}
rescale <- function(x) {
  rng <- range(x)
  (x - rng[1])/(rng[2] - rng[1])
}
```

small test on "rescale" method
```{r}
rescale(1:10)
```

```{r}
rescale(1:3)
```

rescale2 method with *na.rm*
```{r}
rescale2 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  (x - rng[1])/(rng[2] - rng[1])
}
```

Test on rescale2 method
```{r}
rescale2(c(1,2,NA,4,10))
```

rescale3 method
```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
   if(na.rm) {
     rng <-range(x, na.rm=TRUE)
   } else {
     rng <-range(x)
   }
   print("Hello")
   answer <- (x - rng[1]) / (rng[2] - rng[1])
   return(answer)
   print("is it me you are looking for?")
   if(plot) {
      plot(answer, typ="b", lwd=4)
   }
   print("I can see it in ...")
}
```

Test on rescale3
```{r}
rescale3(c(1:6, NA, 10))
```

```{r}
rescale4 <- function(x, na.rm=TRUE, plot=FALSE) {
    rng <-range(x, na.rm=na.rm)
    print("Hello")
   answer <- (x - rng[1]) / (rng[2] - rng[1])
   print("is it me you are looking for?")
   if(plot) {
      plot(answer, typ="b", lwd=4)
   }
   print("I can see it in ...")
   return(answer)
}
```

```{r}
rescale4(c(1:6, NA, 10), plot = TRUE)
```

## Hand-on Section

Question 1B
```{r}
library(bio3d)
s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

Improvement of the above code
```{r}
Analysis <- function(x){
  s <- read.pdb(x)
  s.chainA <- trim.pdb(s, chain = "A", elety = "CA")
  s.b <- s.chainA$atom$b
  plotb3(s.b, sse=s.chainA, typ="l", ylab = "Bfactor")
}
```

Test on the improvement of Analysis method
```{r}
test <- c("4AKE", "1AKE", "1E4Y")
for (i in test){
  Analysis(i)
}
```





