---
title: "BIMM143_class08"
author: "Jiaye Chen"
date: "February 5, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering

```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```

###clustering
####Use the kmeans() function setting k to 2 and nstart=20
####Inspect/print the results
####Q. How many points are in each cluster? 30
####Q. What ‘component’ of your result object details
####      - cluster size? 60
####      - cluster assignment/membership?
####      - cluster center?
####Plot x colored by the kmeans cluster assignment and
####      add cluster centers as blue points

```{r}
km <- kmeans(x, centers = 2, nstart = 20)
```

```{r}
km$size
```

Cluster assignment/ membership and cluster centers

```{r}
plot(x, col=km$cluster)
points(km$centers, col = "blue", pch =15, cex = 2)
```
```{r}
km$totss
```

##Hierarchical clustering in R

```{r}
# First we need to calculate point (dis)similarity
#   as the Euclidean distance between observations
dist_matrix <- dist(x)
# The hclust() function returns a hierarchical
#  clustering model
hc <- hclust(d = dist_matrix)
# the print method is not so useful here
hc

```

```{r}
# Create hierarchical cluster model: hc
hc2 <- hclust(dist(x))
# We can plot the results as a dendrogram
plot(hc2)
```

```{r}
# Draws a dendrogram
plot(hc2)
abline(h=6, col="red") 
cutree(hc2, h=6) # Cut by height h
```

```{r}
# Using different hierarchical clustering methods
d <- dist_matrix
hc.complete <- hclust(d, method="complete")
hc.average  <- hclust(d, method="average")
hc.single   <- hclust(d, method="single")

plot(hc.average)
plot(hc.complete)
plot(hc.single)
```



```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
  matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2),   # c1
  matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
  matrix(c(rnorm(50, mean = 1, sd = 0.3),           # c3
           rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
#         (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```
#### Using dist()
```{r}
plot(dist(x))
```
#### Using hclust()
```{r}
hc3 <- hclust(dist(x))
plot(hc3)
abline(h=2, col="red") 
abline(h=2.5, col = "blue")
cutree(hc3, h=1) # Cut by height h
```

```{r}
gp2 <- cutree(hc3, k=2)
gp3 <- cutree(hc3, k=3)

plot(x, col=gp2)
plot(x, col=gp3)
```

```{r}
plot(hclust(dist(x)))
```

## PCA
```{r}
## You can also download this file from the class website!
mydata <- read.csv("https://tinyurl.com/expression-CSV",row.names=1)
head(t(mydata))
```

```{r}
## lets do PCA
pca <- prcomp(t(mydata), scale=TRUE)
## See what is returned by the prcomp() function
attributes(pca)

## A basic PC1 vs PC2 2-D plot
plot(pca$x[,1], pca$x[,2])
``` 

```{r}
## Variance captured per PC
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
pca.var.per
```
```{r}
barplot(pca.var.per, main="Scree Plot",xlab="Principal Component", ylab="Percent Variation")
```

```{r}
## A vector of colors for wt and ko samples
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
xlab=paste0("PC1 (", pca.var.per[1], "%)"),
ylab=paste0("PC2 (", pca.var.per[2], "%)"))
```

```{r}
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,xlab=paste0("PC1 (", pca.var.per[1], "%)"),ylab=paste0("PC2 (", pca.var.per[2], "%)"))
## Click to identify which sample is which
identify(pca$x[,1], pca$x[,2], labels=colnames(mydata))
```

## Hands-on Section
```{r}
x <- read.csv("UK_foods.csv")

## Number of columns and rows
dim(x)

## Preview the first 6 rows
head(x)

# Note how the minus indexing works
rownames(x) <- x[,1]
x <- x[,-1]
head(x)
dim(x)

## Use the objects in the first column as row names
x <- read.csv("UK_foods.csv", row.names=1)
head(x)

## ploting the data
pairs(x, col=rainbow(10), pch=16)

```

## PCA to rescue

```{r}
# Use the prcomp() PCA function 
pca <- prcomp(t(x))
summary(pca)
```

```{r}
# Plot PC1 vs PC2
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x), col = c("orange", "red", "blue", "darkgreen"))
```


```{r}
## Lets focus on PC1 as it accounts for > 90% of variance 
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```

