---
title: "Class11: Structural Bioinformatics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section 1. The PDB

Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” > “by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project and determine the percentage of structures solved by X-Ray and Electron Microscopy. From the website what proportion of structures are protein? Aim to have a rendered GitHub document with working code that yields your answers.
```{r}
stats <- read.csv("Data Export Summary.csv")
show(stats)
```

```{r}
percent.by.method <- stats$Total/sum(stats$Total)*100
names(percent.by.method) <- stats$Experimental.Method
percent.by.method

```

## BIO3D
```{r}
library(bio3d)
example(plot.bio3d)
```

```{r}
pdb <- read.pdb("1hsg.pdb")
pdb
```

```{r}
attributes(pdb)
```

```{r}
head(pdb$atom)
# Print a subset of $atom data for the first two atoms  
pdb$atom[1:2, c("eleno", "elety", "x","y","z")]
# Note that individual $atom records can also be accessed like this  
pdb$atom$elety[1:2]
# Which allows us to do the following  
plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab ="B-factor")
```
```{r}
# Print a summary of the coordinate data in $xyz
pdb$xyz

```

```{r}
# Select all C-alpha atoms (return their indices)  
ca.inds <- atom.select(pdb, "calpha") 
ca.inds

```

## Write new PDB file
we want to select out the protein and drug only parts of these molecular PDB files
```{r}
prot.inds <- atom.select(pdb,"protein")
pdb$atom[prot.inds$atom]
#write.pdb("1hsg.pdb","ProtOnly.pdb", )
```
```{r}
prot.pdb <- trim.pdb(pdb, prot.inds)
write.pdb(prot.pdb, file = "protein.pdb")
```

Lets do the same thing for our drug
```{r}
lig.inds <- atom.select(pdb,"ligand") 
lig.pdb <- trim.pdb(pdb, lig.inds)
write.pdb(lig.pdb, file = "ligand.pdb")
```

