---
title: "BIMM143_class09"
author: "Jiaye Chen"
date: "February 7, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mini Project

### Preparing the data
```{r}
# Save your input data file to a new 'data' directory
fna.data <- "WisconsinCancer.csv"

# Complete the following code to input the data and store as wisc.df
wisc.df <- read.csv(fna.data)

# View the data
head(wisc.df)
```

```{r}
nrow(wisc.df)
ncol(wisc.df)
```

```{r}
# Convert the column 3-32 into a matrix
wisc.data <- as.matrix(wisc.df[,3:32])
head(wisc.data)
```

```{r}
# Set the row names of wisc.data
row.names(wisc.data) <- wisc.df$id
View(wisc.data)
```

```{r}
View(wisc.df)
```

####Finally, setup a separate new vector called diagnosis to be 1 if a diagnosis is malignant ("M") and 0 otherwise. Note that R coerces TRUE to 1 and FALSE to 0.
```{r}
# Create diagnosis vector by completing the missing code
diagnosis <- as.numeric(wisc.df$diagnosis=="M")
```

#### Q1. How many observations are in this dataset?
#### Q2. How many variables/features in the data are suffixed with _mean?
#### Q3. How many of the observations have a malignant diagnosis?
```{r}
# Q1
nrow(wisc.df)
ncol(wisc.df)
dim(wisc.df)

# Q2
inds <- grep("_mean", colnames(wisc.data))
# the name of these columns
colnames(wisc.data)[inds]

# Q3
sum(diagnosis)
```

## PCA

####Performing PCA
```{r}
# Check column means and standard deviations
colMeans(wisc.data)

apply(wisc.data,2,sd)
```

```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data, scale = TRUE)
# Look at summary of results
summary(wisc.pr)
```
```{r}
#Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?
0.4427
#Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
tracker <- 0
result <- 0
prop_var <- wisc.pr$sdev^2/sum(wisc.pr$sdev^2)
for (i in prop_var){
  if (result < 0.7){
    tracker <- tracker+1
    result <- result+i
  }
  else{
    break
  }
}
tracker
#Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?
tracker <- 0
result <- 0
prop_var <- wisc.pr$sdev^2/sum(wisc.pr$sdev^2)
for (i in prop_var){
  if (result < 0.9){
    tracker <- tracker+1
    result <- result+i
  }
  else{
    break
  }
}
tracker
```
```{r}
biplot(wisc.pr)
```
```{r}
# Scatter plot observations by components 1 and 2
plot(wisc.pr$x[,1],wisc.pr$x[,2], col = diagnosis+1, xlab = "PC1", ylab = "PC2")
```

```{r}
# Repeat for components 1 and 3
plot(wisc.pr$x[,1],wisc.pr$x[,3], col = (diagnosis + 1), 
     xlab = "PC1", ylab = "PC3")
```

```{r}
# Calculate variance of each component
# Variance explained by each principal component: pve
pve <- prop_var

# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", ylab = "Proportion of Variance Explained", ylim = c(0, 1), type = "o")
```

```{r}
# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```
```{r}
# Plot cumulative proportion of variance explained
plot(cumsum(pve), xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```

## Hierachical Clustering

####Hierarchical clustering of case data
```{r}
# Scale the wisc.data data: data.scaled
data.scaled <- scale(wisc.data)
```
```{r}
data.dist <- dist(data.scaled)
```
```{r}
wisc.hclust <- hclust(data.dist, method = "complete")
```
```{r}
plot(wisc.hclust)
abline( h = 19.5, col="red", lty=2)

```

#### Selecting number of clusters
```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, h = 19.5)
table(wisc.hclust.clusters, diagnosis)
```


##Combining methods

####Clustering on PCA results

For clustering we need?
1.Distance matrix
2.Clustering function
3.Cutree

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col= diagnosis+1)
```
```{r}
pc.dist <- dist(wisc.pr$x[,1:7])
pc.hclust <- hclust(pc.dist, method = "ward.D2")
plot(pc.hclust)
```
```{r}
grps <- cutree(pc.hclust, k=2)
table(grps)
```
```{r}
table(grps, diagnosis)
```

```{r}
plot(wisc.pr$x[,1:2], col=grps)
```

```{r}
plot(wisc.pr$x[,1:2], col=diagnosis+1)
```
```{r}
#install.packages("rgl")
library(rgl)
plot3d(wisc.pr$x[,1:3], xlab="PC 1", ylab="PC 2", zlab="PC 3", cex=1.5, size=1, type="s", col=diagnosis+1)
```

## Prediction
We will use the predict() function that will take our PCA model from before and new cancer cell data and project that data onto our PCA space.
```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```
```{r}
plot(wisc.pr$x[,1:2], col=grps)
points(npc[,1], npc[,2], col="blue", pch=16)
```

